NGO analysis

### 2022-07-21 Shiny/Test/Flexdashboard rows.Rmd
Retitled to NGO analysis Plotly caption function. Green/Red scale function. Commenting. Fixing check for invalid value box variables. Commenting out real html function, adding erroneous one for testing.

---
title: "Charity analysis"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    theme: lumen
runtime: shiny
---

@@

  #---- Plots --------------------------------------------------------------------    
  #----CEO compensation-----------------------------------------------------------
  CEO_plot <- reactive({
    htmldata() |>
      # preparing for pivoting
      select(
        year, lederlonn, leder_annen_godtgjorelse
      ) |>
      # Compensation in ordinary numbers
      mutate(
        lederlonn = lederlonn * 1000,
        leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
      ) |>
      # renaming for the purposes of the plot
      rename(
        Year = year,
        Salary = lederlonn,
        `Other compensation` = leder_annen_godtgjorelse
      ) |>
      # pivoting to join the two values to use fill
      pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
      ggplot(aes(Year, value, fill = Compensation)) +
      geom_col() +
      scale_y_continuous(
        # Checking for uniform currency code
        labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
          # Using the currency code if uniform
          label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
          # alternative prefix if multiple currencies in the data
          label_dollar(prefix = "? ")
        )
      ) +
      # pretty colours
      scale_fill_met_d("Troy") +
      labs(
        x = "", y = "Total compensation", title = "CEO compensation"
      )
  })

  # Creating plotly output
  output$CEO_plotly <- renderPlotly(
    ggplotly(CEO_plot()) |>
      layout(
        annotations =
          list(
            x = 1, y = -0.1, text = currency_caption(),
            showarrow = F, xref = "paper", yref = "paper",
            xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
            font = list(size = 11)
          )
      )
  )

  #----Annual result---------------------------------------------
  arsresultat_plot <- reactive({
    htmldata() |>
      ggplot(aes(year, arsresultat, fill = annual_profit)) +
      geom_col() +
      currency_scale() +
      # slightly nicer colours + ensuring positive values are always green
      scale_fill_met_d("Tsimshian",
        direction =
          ifelse(mean(htmldata()$annual_profit,
            na.rm = TRUE
          ) == 1, -1, 1)
      ) +
      theme(legend.position = "none") +
      labs(
        x = "", y = "", title = "Annual result (in thousands)"
      )
  })

  output$arsresultat_plotly <- renderPlotly(
    ggplotly(arsresultat_plot()) |>
      layout(
        annotations =
          list(
            x = 1, y = -0.1, text = currency_caption(),
            showarrow = F, xref = "paper", yref = "paper",
            xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
            font = list(size = 11)
          )
      )
  )
  #----Balance--------------------------------------------------------------------    
  balance_plot <- reactive({
    htmldata() |>
      ggplot(aes(year, sum_egenkapital_og_gjeld, fill = balance_positive)) +
      geom_col() +
      currency_scale() +
      # slightly nicer colours + ensuring positive values are always green
      scale_fill_met_d("Tsimshian",
        direction = ifelse(
          mean(htmldata()$balance_positive,
            na.rm = TRUE
          ) == 1, -1, 1
        )
      ) +
      theme(legend.position = "none") +
      labs(
        x = "", y = "", title = "Balance (in thousands)",
      )
  })

  output$balance_plotly <- renderPlotly(
    ggplotly(balance_plot()) |>
      layout(
        annotations =
          list(
            x = 1, y = -0.1, text = currency_caption(),
            showarrow = F, xref = "paper", yref = "paper",
            xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
            font = list(size = 11)
          )
      )
  )

  #----Expenses---------------------------------------
  expenses_plot <- reactive({
    exceldata() |>
      transmute(
        Year = as.factor(year),
        Other = expenses_other,
        Administrative = expenses_administrative,
        Fundraising = expenses_fundraising,
        `Program services` = expenses_program_services
      ) |>
      pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
      mutate(
        Expenses = ordered(Expenses,
          levels =
            c(
              "Program services", "Administrative",
              "Fundraising", "Other"
            )
        )
      ) |>
      ggplot(aes(Year, value, fill = Expenses)) +
      geom_col() +
      coord_flip() +
      currency_scale() +
      scale_fill_met_d("Troy") +
      labs(x = "", y = "", title = "Expenses (in thousands)")
  })

  output$expenses_plotly <- renderPlotly(
    ggplotly(expenses_plot()) |>
      layout(
        annotations =
          list(
            x = 1, y = -0.1, text = currency_caption(),
            showarrow = F, xref = "paper", yref = "paper",
            xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
            font = list(size = 11)
          )
      )
  )

@@

  # Reactive function to create currency caption
  currency_caption <- reactive({
    if (length(unique(htmldata()$valutakode)) == 1) {
      ""
    } else {
      "Note: not all values are denominated in the same currency"
    }
  })
})

@@

```{r non-reactive functions, include=FALSE}
# Function to read html file
read_proff8 <- function(file, start, end) {
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame") |>
    html_table() |>
    select( # removing empty column
      REGNSKAPSPERIODE:all_of(start)
    ) |> # using all_of()
    mutate( # removing whitespace
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    filter( # removing duplicate table - no idea why it's necessary
      !row_number() > 176
    ) |>
    filter( # removing duplicate rows
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    pivot_longer( # tidying data
      all_of(end):all_of(start),
      names_to = "year"
    ) |> # using all_of()
    mutate( # changing to real NAs and turning years into numbers
      value = na_if(value, "-"),
      year = as.integer(year),
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    fill( # filling the currency column
      valutakode,
      .direction = "updown"
    ) |>
    filter( # removing dates and redundant currency
      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
        REGNSKAPSPERIODE != "Valutakode"
    ) |>
    mutate( # removing whitespace in numbers
      value = str_replace_all(value, "\\s", ""),
      value = as.numeric(value),
      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
    ) |>
    distinct() |> # removing any remaining duplicate rows
    pivot_wider( # pivoting
      names_from = REGNSKAPSPERIODE, values_from = value
    ) |>
    clean_names() |> # tidying names
    select(
      -lederlonn_i_hele_1000,
      -resultatregnskap_i_hele_1000,
      -balanseregnskap_i_hele_1000
    ) |> # removing headings
    rename( # making it clear what lonn refers to
      lederlonn = lonn
    ) |>
    # Creating new variables for plots
    mutate(
      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    ) |>
    arrange(desc(year)) # arranging by year
}

@@

# Function to turn KPI values into NA or percentage string
KPI_to_string <- function(variable){
  if (variable == 0 | is.na(variable)){
    NA
  } else {
    str_c(format(variable, digits = 3), "%")
  }
}

>>>>

---
title: "NGO analysis"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    theme: lumen
runtime: shiny
---

@@

  #---- Plots --------------------------------------------------------------------    
  #----CEO compensation-----------------------------------------------------------
  CEO_plot <- reactive({
    htmldata() |>
      # preparing for pivoting
      select(
        year, lederlonn, leder_annen_godtgjorelse
      ) |>
      # Compensation in ordinary numbers
      mutate(
        lederlonn = lederlonn * 1000,
        leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
      ) |>
      # renaming for the purposes of the plot
      rename(
        Year = year,
        Salary = lederlonn,
        `Other compensation` = leder_annen_godtgjorelse
      ) |>
      # pivoting to join the two values to use fill
      pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
      ggplot(aes(Year, value, fill = Compensation)) +
      geom_col() +
      currency_scale() +
      # pretty colours
      scale_fill_met_d("Troy") +
      labs(
        x = "", y = "Total compensation", title = "CEO compensation"
      )
  })

  # Creating plotly output
  output$CEO_plotly <- renderPlotly(
    ggplotly(CEO_plot()) |>
      layout(
        annotations =
          plotly_caption()
      )
  )

  #----Annual result---------------------------------------------
  arsresultat_plot <- reactive({
    htmldata() |>
      ggplot(aes(year, arsresultat, fill = annual_profit)) +
      geom_col() +
      currency_scale() +
      scale_fill_green() +
      theme(legend.position = "none") +
      labs(
        x = "", y = "", title = "Annual result (in thousands)"
      )
  })

  output$arsresultat_plotly <- renderPlotly(
    ggplotly(arsresultat_plot()) |>
      layout(
        annotations =
          plotly_caption()
      )
  )
  #----Balance--------------------------------------------------------------------    
  balance_plot <- reactive({
    htmldata() |>
      ggplot(aes(year, sum_egenkapital_og_gjeld, fill = balance_positive)) +
      geom_col() +
      currency_scale() +
      scale_fill_green() +
      theme(legend.position = "none") +
      labs(
        x = "", y = "", title = "Balance (in thousands)",
      )
  })

  output$balance_plotly <- renderPlotly(
    ggplotly(balance_plot()) |>
      layout(
        annotations =
          plotly_caption()
      )
  )

  #----Expenses---------------------------------------
  expenses_plot <- reactive({
    exceldata() |>
      # Renaming for the purposes of the plot + dropping other variables
      transmute(
        Year = as.factor(year),
        Other = expenses_other,
        Administrative = expenses_administrative,
        Fundraising = expenses_fundraising,
        `Program services` = expenses_program_services
      ) |>
      # Pivoting to use fill
      pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
      ggplot(aes(Year, value, fill = Expenses)) +
      geom_col() +
      # Flipping, because the Excel file only contains two years atm
      coord_flip() +
      currency_scale() +
      scale_fill_met_d("Troy") +
      labs(x = "", y = "", title = "Expenses (in thousands)")
  })

  output$expenses_plotly <- renderPlotly(
    ggplotly(expenses_plot()) |>
      layout(
        annotations =
          plotly_caption()
      )
  )

@@

  # Reactive function to create currency caption
  currency_caption <- reactive({
    # Checking for uniform currency code
    if (length(unique(htmldata()$valutakode)) == 1) {
      # Empty caption if uniform
      ""
      # Warning if multiple currencies
    } else {
      "Note: not all values are denominated in the same currency"
    }
  })

  # Reactive function to create plotly caption
  # (ggplot captions and subtitles are lost in the conversion)
  # (shamelessly stolen from the Internet, with minor modifications)
  plotly_caption <- reactive({
    list(
      x = 1, y = -0.1, text = currency_caption(),
      showarrow = F, xref = "paper", yref = "paper",
      xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
      font = list(size = 11)
    )
  })

  # Reactive function with slightly nicer colours + ensuring positive values are always green
  scale_fill_green <- reactive({
    scale_fill_met_d("Tsimshian",
      direction = ifelse(
        mean(htmldata()$balance_positive,
          na.rm = TRUE
        ) == 1, -1, 1
      )
    )
  })
})

@@

```{r non-reactive functions, include=FALSE}
# Function to read html file
# read_proff8 <- function(file, start, end) {
#   read_html(file, encoding = "UTF-8") |>
#     html_node("#inner-frame") |>
#     html_table() |>
#     select( # removing empty column
#       REGNSKAPSPERIODE:all_of(start)
#     ) |> # using all_of()
#     mutate( # removing whitespace
#       REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
#     ) |>
#     filter( # removing duplicate table - no idea why it's necessary
#       !row_number() > 176
#     ) |>
#     filter( # removing duplicate rows
#       !grepl("Lukk", REGNSKAPSPERIODE)
#     ) |>
#     pivot_longer( # tidying data
#       all_of(end):all_of(start),
#       names_to = "year"
#     ) |> # using all_of()
#     mutate( # changing to real NAs and turning years into numbers
#       value = na_if(value, "-"),
#       year = as.integer(year),
#       valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
#     ) |>
#     fill( # filling the currency column
#       valutakode,
#       .direction = "updown"
#     ) |>
#     filter( # removing dates and redundant currency
#       REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
#         REGNSKAPSPERIODE != "Valutakode"
#     ) |>
#     mutate( # removing whitespace in numbers
#       value = str_replace_all(value, "\\s", ""),
#       value = as.numeric(value),
#       value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
#     ) |>
#     distinct() |> # removing any remaining duplicate rows
#     pivot_wider( # pivoting
#       names_from = REGNSKAPSPERIODE, values_from = value
#     ) |>
#     clean_names() |> # tidying names
#     select(
#       -lederlonn_i_hele_1000,
#       -resultatregnskap_i_hele_1000,
#       -balanseregnskap_i_hele_1000
#     ) |> # removing headings
#     rename( # making it clear what lonn refers to
#       lederlonn = lonn
#     ) |>
#     # Creating new variables for plots
#     mutate(
#       annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
#       balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
#     ) |>
#     arrange(desc(year)) # arranging by year
# }

# Test version of function - multiple currencies
read_proff8 <- function(file, start, end) {
  read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame") |>
    html_table() |>
    select( # removing empty column
      REGNSKAPSPERIODE:all_of(start)
    ) |> # using all_of()
    mutate( # removing whitespace
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    filter( # removing duplicate table - no idea why it's necessary
      !row_number() > 176
    ) |>
    filter( # removing duplicate rows
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    pivot_longer( # tidying data
      all_of(end):all_of(start),
      names_to = "year"
    ) |> # using all_of()
    mutate( # changing to real NAs and turning years into numbers
      value = na_if(value, "-"),
      year = as.integer(year),
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    fill( # filling the currency column
      valutakode,
      .direction = "updown"
    ) |>
    filter( # removing dates and redundant currency
      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
        REGNSKAPSPERIODE != "Valutakode"
    ) |>
    mutate( # removing whitespace in numbers
      value = str_replace_all(value, "\\s", ""),
      value = as.numeric(value),
      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
    ) |>
    distinct() |> # removing any remaining duplicate rows
    pivot_wider( # pivoting
      names_from = REGNSKAPSPERIODE, values_from = value
    ) |>
    clean_names() |> # tidying names
    select(
      -lederlonn_i_hele_1000,
      -resultatregnskap_i_hele_1000,
      -balanseregnskap_i_hele_1000
    ) |> # removing headings
    rename( # making it clear what lonn refers to
      lederlonn = lonn
    ) |>
    # Creating new variables for plots
    mutate(
      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    ) |>
    # bad test - REMOVE
    mutate(
    valutakode = ifelse(year == 2019, "USD", valutakode)
  ) |>
    arrange(desc(year)) # arranging by year
}

@@

# Function to turn KPI values into NA or percentage string
KPI_to_string <- function(variable){
  # Returning NA if the KPI is Inf (division by zero) or NA, to avoid
  # layout-breaking errors, otherwise formatting
  if (is.na(variable)){
    NA
  } else if (variable == Inf){
    NA
  } else {
    # Shortening (lengthening) the value, adding %
    str_c(format(variable, digits = 3), "%")
  }
}
